### 2021-02-02 manuscript/abstract.tex
Apply GJI style to manuscript and refactor it (#115)
Refactor manuscript by splitting it into several files. Create
a manuscript-gji.tex file that creates a version of the manuscript following
the GJI style. Both manuscript.tex and manuscript-gji.tex don't contain
actual content, just compiles the content of the other files. Move every .tex
file created by the notebooks to manuscript/results/. Comment out the
definition of \nl command on gji_extra.sty: it's already being defined by
algorithm2e. Update the Makefile to the latest changes, add targets for
building the GJI version of the manuscrip

The equivalent source technique is a powerful and widely used method for
processing gravity and magnetic data.  Nevertheless, its major
drawback is the large computational cost in terms of processing time and
computer memory.
We present two techniques for reducing the computational cost of equivalent
source processing: block-averaging source locations and the
gradient-boosted equivalent source algorithm.
By using block-averaging, we reduce the number of source coefficients that
must be estimated while retaining the minimum desired resolution in the final
processed data.
With the gradient boosting method, we estimate the layer coefficients in
small batches along overlapping windows, allowing us to reduce the computer
memory requirements arbitrarily to conform to the constraints of the
available hardware.
We show that the combination of block-averaging and gradient-boosted
equivalent sources are capable of producing accurate interpolations through
tests against synthetic data.
Moreover, we demonstrate the feasibility of our method by gridding a gravity
dataset covering Australia with over 1.7 million observations using a modest
personal computer.