### 2021-02-02 manuscript/manuscript.tex
Configure fontsize and geometry of manuscript

\documentclass[twocolumn,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[longend,linesnumbered,ruled]{algorithm2e}
\usepackage[pdftex,colorlinks=true]{hyperref}
\usepackage{fancyhdr}

% Import packages useful for creating the parameters tables
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{setspace}

% Import natbib and doi packages
\usepackage[round,authoryear,sort]{natbib}
\usepackage{doi} % show dois as links on references

% Define style for urls
\urlstyle{same}

% Configure column separation and page geometry
\usepackage[hmargin=2cm, vmargin=3cm]{geometry}
\setlength{\columnsep}{2\columnsep}

% Import variables
\input{variables.tex}

% Define title, authors, affiliations and DOI
% ===========================================
\newcommand{\Author}{%
    \SolerShort{} and \UiedaShort{}
}
\newcommand{\AuthorAffil}{%
    {\large
        \Soler$^{1,2}$ and \Uieda$^{3}$
    }
    \\[0.4cm]
    {\small $^{1}$\AffiliationA} \\
    {\small $^{2}$\AffiliationB} \\
    {\small $^{3}$\AffiliationC} \\
}

% Configure header and hypersetup
% ===============================
\pagestyle{fancy}
\fancyhf{}
\lhead{
    \fontsize{9pt}{12pt}\selectfont
    \Author{} (\the\year{}). \DOI{}
}
\rhead{\fontsize{9pt}{12pt}\selectfont \thepage}
\renewcommand{\headrulewidth}{0pt}
\hypersetup{
    allcolors=blue,
    pdftitle={\Title},
    pdfauthor={\Author},
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{\Title}
\author{\AuthorAffil}
\date{
    \normalsize
    \today
}
\maketitle

\noindent
\textbf{Keywords:}
\input{keywords.tex}

\begin{abstract}
    \input{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{content.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{plainnat}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{appendix.tex}

\end{document}